---
title: "GEOG 415 — Lab 3 SOLUTION / TA answer key"
subtitle: "Exploratory analysis of the analysis-ready dataset"
format:
  html:
    toc: true
execute:
  eval: false     # set TRUE once data_raw/tn_tract_analysis.csv + tn_tracts.gpkg are present
  echo: true
---

<!--
INSTRUCTOR / TA ONLY — not distributed to students.

Reference run for Lab 3. Uses the provided analysis-ready dataset
(tn_tract_analysis.csv, built by data/analytic.R and shipped in the Lab 3
increment) joined to tn_tracts.gpkg from Lab 1. Run from inside a project that
has both files in data_raw/.
-->

```{r setup}
library(tidyverse)
library(sf)

d      <- readr::read_csv("data_raw/tn_tract_analysis.csv", show_col_types = FALSE)
tracts <- sf::st_read("data_raw/tn_tracts.gpkg", quiet = TRUE)
tj     <- tracts |> left_join(d, by = "GEOID")

candidates <- c("poverty_rate","novehicle_rate","uninsured_pct","age65_pct","disability_pct")
```

## Distributions

```{r dist}
d |>
  select(all_of(candidates)) |>
  pivot_longer(everything(), names_to = "indicator", values_to = "value") |>
  ggplot(aes(value)) +
  geom_histogram(bins = 30) +
  facet_wrap(~ indicator, scales = "free") +
  theme_minimal()
```

## Correlations / redundancy

```{r cor}
d |> select(all_of(candidates)) |> cor(use = "pairwise.complete.obs") |> round(2)
```

Expect poverty, uninsured, and no-vehicle to be moderately-to-strongly related
(shared deprivation), while age65 and disability track each other and are
partly distinct from the economic cluster — the discussion students should have.

## Maps

```{r maps}
ggplot(tj) + geom_sf(aes(fill = poverty_rate), color = NA) +
  scale_fill_viridis_c(na.value = "grey90") + theme_minimal()
```

## Context: metro vs non-metro

```{r context}
d <- d |>
  mutate(metro = if_else(county_geoid %in% c("47037","47157","47093","47065"),
                         "metro", "non-metro"))
d |> group_by(metro) |>
  summarise(cor_pov_unins = cor(poverty_rate, uninsured_pct, use = "pairwise.complete.obs"),
            n = n())
```

## Scale: tract vs county (the MAUP point)

```{r maup}
county_right <- d |>
  group_by(county_geoid) |>
  summarise(poverty_count = sum(poverty_count, na.rm = TRUE),
            poverty_universe = sum(poverty_universe, na.rm = TRUE)) |>
  mutate(poverty_rate_county = poverty_count / poverty_universe)

county_wrong <- d |>
  group_by(county_geoid) |>
  summarise(poverty_rate_avg = mean(poverty_rate, na.rm = TRUE))

# How far the wrong (mean-of-rates) method drifts from the right one:
county_right |>
  left_join(county_wrong, by = "county_geoid") |>
  mutate(diff = poverty_rate_avg - poverty_rate_county) |>
  arrange(desc(abs(diff)))
```

## Outliers as places

```{r outliers}
d |> slice_max(poverty_rate, n = 5) |> select(GEOID, tract_name, total_population, poverty_rate)
d |> filter(total_population < 200) |> arrange(desc(poverty_rate)) |>
  select(GEOID, tract_name, total_population, poverty_rate)
```

## Grading notes

- **Aggregation:** full credit requires summing numerator/denominator, not
  averaging rates; strong answers show the two differ and explain why
  (population weighting). Bonus for noting the subject-table percents
  (`uninsured_pct`, `age65_pct`, `disability_pct`) *cannot* be re-aggregated.
- **Redundancy:** expect an explicit call on at least one correlated pair.
- **Fitness:** decisions must argue about measuring *primary-care need*, not just
  statistical correlation.
- **Places, not rows:** outliers discussed as real tracts (small population,
  group quarters) rather than anonymous extreme values.
