---
title: "GEOG 415 — Data Provenance & Measurement Audit"
subtitle: "Community Primary Care Priority Rating — Module 1"
author: "Team: ___"
date: today
format:
  html:
    toc: true
    number-sections: true
  pdf:
    toc: true
    number-sections: true
execute:
  echo: true
  warning: false
  message: false
---

<!--
This is your Module 1 deliverable. It is a Quarto document on purpose: the
tables and maps below are generated from your code, so rendering it is proof
your work is reproducible. Fill in every section marked TODO. The poverty rate
is worked out for you as a model — write the other variables to the same
standard.

Do NOT put your Census API key in this file.
-->

```{r setup}
library(tidyverse)
library(sf)

# Read the data you produced in Lab 2. Adjust paths/names to match your scripts.
acs    <- readr::read_csv("data_processed/analytic_v1.csv", show_col_types = FALSE)
tracts <- sf::st_read("data_raw/tn_tracts.gpkg", quiet = TRUE)
```

# Purpose and intended use

<!-- TODO: 1–2 paragraphs. What decision could this rating inform? State plainly
what it is FOR and what it is NOT for (e.g., not for ranking individual people,
not a measure of health outcomes). -->

# Source inventory

For every source: agency, dataset, vintage, access date, endpoint/URL, coverage.

| Source | Dataset | Vintage | Accessed | Endpoint / URL | Coverage |
|--------|---------|---------|----------|----------------|----------|
| U.S. Census Bureau | ACS 5-year Detailed & Subject Tables | `____` | `____` | `api.census.gov/.../acs5` (via tidycensus) | TN census tracts |
| HRSA | Health Center service-delivery sites | `____` | `____` | `____ (paste exact download URL)` | U.S. (filtered to TN) |
| U.S. Census Bureau | TIGER/Line tracts (from Lab 1) | `____` | `____` | `www2.census.gov/geo/tiger/...` | TN census tracts |

<!-- TODO: fill in every blank with the real value you used. -->

# Data dictionary

<!-- TODO: one row per variable you KEPT. Include the plain-language definition,
units, the denominator (for rates), and a margin-of-error note where relevant. -->

| Variable | Definition (plain language) | Units | Denominator | MOE? |
|----------|-----------------------------|-------|-------------|------|
| `poverty_rate` | Share of the population (for whom poverty status is determined) with income below the poverty level | proportion 0–1 | `poverty_universe` (B17001_001) | inherits ACS MOE on numerator & denominator |
| `novehicle_rate` | `____` | `____` | `____` | `____` |
| `uninsured_pct` | `____` | `____` | `____` | `____` |
| `age65_pct` | `____` | `____` | `____` | `____` |
| `disability_pct` | `____` | `____` | `____` | `____` |

# Variable explanations

## Poverty rate (worked example — read this, then do the others)

**What it is.** The poverty rate here is `poverty_count / poverty_universe` from
ACS table B17001: the number of people with income below the federal poverty
level, divided by the population for whom poverty status is determined.

**What it is not.** It is not the share of *all* residents in poverty — the
denominator excludes people for whom poverty status is not determined (for
example, most people living in institutional group quarters, and children in
some living arrangements). It is not a measure of hardship depth: a tract just
below and just above the line look identical in the count, and the federal
threshold does not adjust for local cost of living.

**Uncertainty.** As a 5-year ACS estimate it carries a margin of error that can
be large in small-population tracts; the rate inherits uncertainty from both the
numerator and the denominator. Treat a single tract's value as an estimate, not
a fact.

**Fitness for our purpose.** Poverty is a reasonable *component* of community
need for primary-care priority, but it is not the same construct — a low-poverty
tract can still have poor access, and poverty correlates with other things
(rurality, age) we must not double-count. Use it as one input, not the answer.

```{r poverty-summary}
acs |>
  summarise(
    n_tracts = n(),
    n_missing = sum(is.na(poverty_rate)),
    median = median(poverty_rate, na.rm = TRUE),
    min = min(poverty_rate, na.rm = TRUE),
    max = max(poverty_rate, na.rm = TRUE)
  )
```

## No-vehicle rate

<!-- TODO: write it to the same standard: what it is, what it is not,
uncertainty, fitness for purpose. Add a small summary chunk like the one above. -->

## Uninsured (%)

<!-- TODO. Note: this came pre-computed as a percent from a subject table. What
denominator is hidden inside it, and why does that matter here? -->

## Age 65+ (%) and Disability (%)

<!-- TODO. For each: what it measures, what it does not, fitness for purpose. -->

# Descriptive maps

<!-- TODO: include your saved maps (or regenerate them here). For each, one
sentence on what to notice and one on what the map might wrongly suggest. -->

# Weaknesses and fitness for purpose

A structured audit. For each item, answer: **What does it measure? What does it
*not*? Who is mismeasured, miscounted, or missing? Is it fit for deciding
primary-care priority?**

- **ACS estimates in small tracts.** <!-- TODO -->
- **Rates vs. hidden denominators (the subject-table percents).** <!-- TODO -->
- **HRSA points.** Is a "site" a place where care is delivered, an
  administrative address, or a mailing address? Do all rows have coordinates?
  What does a dot *not* tell you about capacity, hours, cost, or who is served?
  <!-- TODO -->
- **Point vs. areal representation.** <!-- TODO: what does overlaying points on
  tract averages tempt a reader to conclude that the data cannot support? -->
- **Whom does this data make invisible?** <!-- TODO -->

# Reproducibility note

<!-- TODO: list your scripts in run order and confirm a fresh-session rerun
rebuilds the processed data, the maps, and this document from the raw downloads,
with no manual steps and no committed API key. -->
