---
title: "GEOG 415 — Lab 2 SOLUTION / TA answer key"
subtitle: "Reference run: acquire ACS + HRSA, describe, audit"
format:
  html:
    toc: true
    code-fold: false
execute:
  eval: false        # set TRUE once a key is set and the HRSA URL is pinned
  echo: true
  warning: true
  message: true
---

<!--
INSTRUCTOR / TA ONLY — do not distribute to students.

This is the reference version of what students produce in Lab 2. It runs the
whole pipeline end to end so a TA can (a) check student work against a known-good
result and (b) confirm the lab still works before class each term.

Before running:
  1. Set CENSUS_API_KEY in ~/.Renviron (see labs/lab02/templates/Renviron.example).
  2. Pin YEAR to the class vintage and VERIFY the subject-table codes with
     load_variables() — subject-table cells (S....) move between vintages.
  3. Paste the exact HRSA CSV download URL into HRSA_CSV_URL.
  4. Flip eval to TRUE in the YAML above.
-->

```{r setup}
library(tidyverse)
library(tidycensus)
library(sf)

YEAR         <- 2022                     # <-- pin to class vintage
HRSA_CSV_URL <- "PASTE_EXACT_HRSA_URL"   # <-- pin the exact download URL
STATE        <- "47"                     # Tennessee

census_api_key(Sys.getenv("CENSUS_API_KEY"))
stopifnot(nzchar(Sys.getenv("CENSUS_API_KEY")))

# Create the project output folders if they do not already exist, so this
# reference run works from a clean checkout.
for (d in c("data_raw", "data_processed", "figures")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

## Verify variable codes for the vintage

```{r catalogue}
vars <- load_variables(YEAR, "acs5", cache = TRUE)
subj <- load_variables(YEAR, "acs5/subject", cache = TRUE)
# Confirm the S.... codes below against `subj` before trusting them.
```

## Pull ACS

Two count-based indicators (students compute the rate) and three subject-table
percents. Codes below are the intended set; re-verify the subject codes for YEAR.

```{r acs}
acs_vars <- c(
  total_pop      = "B01003_001",
  pov_num        = "B17001_002",
  pov_denom      = "B17001_001",
  novehicle_num  = "B08201_002",
  hh_denom       = "B08201_001",
  uninsured_pct  = "S2701_C05_001",   # percent uninsured — VERIFY for YEAR
  age65_pct      = "S0101_C02_030",   # percent 65 and over — VERIFY for YEAR
  disability_pct = "S1810_C03_001"    # percent with a disability — VERIFY for YEAR
)

acs_raw <- get_acs(
  geography = "tract", state = STATE, year = YEAR,
  survey = "acs5", variables = acs_vars, output = "wide"
)

readr::write_csv(acs_raw, "data_raw/acs_tn_tract_raw.csv")

acs <- acs_raw |>
  mutate(
    poverty_rate   = pov_numE / pov_denomE,
    novehicle_rate = novehicle_numE / hh_denomE,
    uninsured_pct  = uninsured_pctE,
    age65_pct      = age65_pctE,
    disability_pct = disability_pctE
  ) |>
  select(GEOID, poverty_rate, novehicle_rate, uninsured_pct, age65_pct, disability_pct)

readr::write_csv(acs, "data_processed/analytic_v1.csv")
```

## Pull HRSA and build points

```{r hrsa}
download.file(HRSA_CSV_URL, "data_raw/hrsa_health_centers.csv", mode = "wb")
hrsa <- readr::read_csv("data_raw/hrsa_health_centers.csv")

# Column names vary by download — inspect and adjust these three.
hrsa_tn <- hrsa |>
  filter(.data[["State"]] == "TN") |>
  filter(!is.na(.data[["Longitude"]]), !is.na(.data[["Latitude"]])) |>
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)
```

## Join and map

```{r maps}
tracts <- st_read("data_raw/tn_tracts.gpkg", quiet = TRUE)   # from Lab 1
tj <- tracts |> left_join(acs, by = "GEOID")

ggplot(tj) +
  geom_sf(aes(fill = poverty_rate), color = NA) +
  scale_fill_viridis_c(na.value = "grey90") +
  labs(title = "Poverty rate by tract, TN") + theme_minimal()
ggsave("figures/poverty_rate_map.png", width = 8, height = 6, dpi = 300)

ggplot() +
  geom_sf(data = tracts, fill = "grey95", color = "white") +
  geom_sf(data = hrsa_tn, size = 0.8, alpha = 0.7) +
  labs(title = "HRSA sites, TN") + theme_minimal()
ggsave("figures/hrsa_points_map.png", width = 8, height = 6, dpi = 300)
```

## Quick checks a TA can grade against

```{r checks}
list(
  n_tracts      = nrow(acs),
  pov_in_range  = all(acs$poverty_rate >= 0 & acs$poverty_rate <= 1, na.rm = TRUE),
  pov_missing   = sum(is.na(acs$poverty_rate)),
  hrsa_tn_sites = nrow(hrsa_tn)
)
```

## Grading notes

- **Reproducibility:** raw pulls saved unedited; rates computed in script; API key
  absent from all submitted files; HRSA URL + access date recorded.
- **Definitions:** poverty & no-vehicle built from num/denom; the three percents
  are pre-computed — students should name the hidden denominator.
- **Fitness for purpose:** expect specific points on ACS MOE in small tracts,
  point-vs-areal for HRSA, and who is missing (group quarters, uncounted).
